### 2016-11-17 capability-machine.tex
Added getp and getl

%Functions
\newcommand{\plainfun}[2]{
  \ifthenelse{\equal{#2}{}}
             {\mathit{#1}}
             {\mathit{#1}(#2)}
}
\newcommand{\decode}{\plainfun{decode}{}}
\newcommand{\encode}{\plainfun{encode}{}}
\newcommand{\encodePerm}{\plainfun{encodePerm}{}}
\newcommand{\decodePerm}{\plainfun{decodePerm}{}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}

@@

%Instructions
%No arguments
\newcommand{\zinstr}[1]{\mathtt{#1}}
\newcommand{\fail}{\zinstr{fail}}
\newcommand{\halt}{\zinstr{halt}}
%One argument
\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
%Two arguments
\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
%Three arguments
\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}

@@

\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
$$\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\memheap&\in& \Heaps \\
\memreg &\in& \Regs \\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
n       &\in& \ints\\
\end{array}$$
Further definitions:
$$\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& 
             \jmp{\lv} \mid 
             \jnz{\lv}{\rv} \mid
             \move{\lv}{\rv} \mid 
             \load{\lv}{\hv} \mid 
             \store{\hv}{\rv} \mid  \\
       &   & \plus{\lv}{\rv}{\rv} \mid 
             \lea{\lv}{\rv} \mid 
             \restricttwo{\lv}{\rv} \mid 
             \subseg{\lv}{\rv}{\rv} \mid  \\
       &   & \isptr{\lv}{\rv} \mid 
             \getp{\lv}{\lv} \mid 
             \getb{\lv}{\lv} \mid
             \gete{\lv}{\lv} \mid
             \geta{\lv}{\lv} \mid \\
       &   & \fail \mid
             \halt \mid 
\end{array}$$

@@

Assume an $\encodePerm$ function that encodes a permission as an integer:

@@

\encodePerm &: (\Perms \times \Globals) \fun \ints

@@

Further, assume an inverse function, $\decodePerm$, that decodes permissions

@@

  \decodePerm : \ints \fun (\Perms \times \Globals)

@@

                                                   \quad\text{and $\var{newPermPair} = \decodePerm(n)$}\\

@@

                            \stdUpdatePc{\updateReg{r_1}{\perm}} & \text{if $\memreg{r_2} = \stdcap$} \\

@@

%getl?

@@

\subsection{Another simple well-bracketedness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
\begin{verbatim}
let f = fun adv =>
          let l = 0 in
          let x = 1 in
          adv(l);
          assert(x == 1);
          x := 2;
          adv(0);
          return
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
let h = fun adv =>
          let x = 1 in
          let g = fun adv =>
                    x := 0;
                    adv(g);
                    x := 1;
                    adv(g);
                    assert(x == 1) in
          g(g)
\end{verbatim}
In Dreyer's example, the issue arises as the context can provide the result of evaluating $\mathtt{e_1}$ (the equivalent here is \texttt{g})) to \texttt{f} (the equivalent here is \texttt{adv}). In our system, the context cannot capture this, so in the above program, \texttt{adv} is given explicit access to \texttt{g}. The lemma we would want to prove here is that the assertion does not fail.

@@

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                             \begin{aligned}[t]
                               \{ \npair{(\reg,\hs)} \mid & \forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                                                          & \quad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'} \}
                             \end{aligned}
\end{align*}

>>>>

%Functions
\newcommand{\plainfun}[2]{
  \ifthenelse{\equal{#2}{}}
             {\mathit{#1}}
             {\mathit{#1}(#2)}
}
\newcommand{\decode}{\plainfun{decode}{}}
\newcommand{\encode}{\plainfun{encode}{}}
\newcommand{\encodePerm}{\mathit{encodePerm}}
\newcommand{\encodePermPair}{\plainfun{encodePermPair}{}}
\newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
\newcommand{\decodePermPair}{\plainfun{decodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}

@@

%Instructions
%No arguments
\newcommand{\zinstr}[1]{\mathtt{#1}}
\newcommand{\fail}{\zinstr{fail}}
\newcommand{\halt}{\zinstr{halt}}
%One argument
\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
%Two arguments
\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
\newcommand{\getl}[2]{\twoinstr{getl}{#1}{#2}}
\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
%Three arguments
\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}

@@

\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
\[
\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\memheap&\in& \Heaps \\
\memreg &\in& \Regs \\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
n       &\in& \ints\\
\end{array}
\]
Further definitions:
\[
\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& 
             \jmp{\lv} \mid 
             \jnz{\lv}{\rv} \mid
             \move{\lv}{\rv} \mid 
             \load{\lv}{\hv} \mid 
             \store{\hv}{\rv} \mid  \\
       &   & \plus{\lv}{\rv}{\rv} \mid 
             \lea{\lv}{\rv} \mid 
             \restricttwo{\lv}{\rv} \mid 
             \subseg{\lv}{\rv}{\rv} \mid  \\
       &   & \isptr{\lv}{\rv} \mid 
             \getp{\lv}{\lv} \mid 
             \getl{\lv}{\lv} \mid 
             \getb{\lv}{\lv} \mid
             \gete{\lv}{\lv} \mid
             \geta{\lv}{\lv} \mid \\
       &   & \fail \mid
             \halt \mid 
\end{array}
\]

@@

Assume an $\encodePerm$, $\encodeLoc$, and $\encodePermPair$ function that encodes a permissions, locality, and permission pair, respectively, as an integer:

@@

\encodePerm &: \Perms \fun \ints \\
\encodeLoc &: \Globals \fun \ints \\
\encodePermPair &: (\Perms \times \Globals) \fun \ints \\

@@

Further, assume an inverse function, $\decodePermPair$, that decodes permissions

@@

  \decodePermPair : \ints \fun (\Perms \times \Globals)

@@

                                                   \quad\text{and $\var{newPermPair} = \decodePermPair(n)$}\\

@@

                            \stdUpdatePc{\updateReg{r_1}{\encodePerm(\perm)}} &

@@

  \sem{\getl{\refreg{r_1}}{\refreg{r_2}}} & = 
                            \stdUpdatePc{\updateReg{r_1}{\encodeLoc(\gl)}} &

@@

\subsection{Another simple well-bracketedness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          let l = 0 in
          adv(l);
          assert(x == 1);
          x := 2;
          adv(0);
          return
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
let h = fun adv =>
          let x = 1 in
          let g = fun g' =>
                    x := 1;
                    adv(g');
                    x := 0;
                    adv();
                    assert(x == 0) in
          g(g);
          return
\end{verbatim}
In Dreyer's example, the issue arises as the context can provide the result of evaluating $\mathtt{e_1}$ (the equivalent here is \texttt{g})) to \texttt{f} (the equivalent here is \texttt{adv}). In our system, the context cannot capture this, so in the above program, \texttt{adv} is given explicit access to \texttt{g}. The lemma we would want to prove here is that the assertion does not fail.

@@

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                               \{ \npair{(\reg,\hs)} \mid
                             \begin{aligned}[t]
                               & \forall \heap_f, \heap', i \leq n \ldotp \\
                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                               & \qquad \Rightarrow
                               \begin{aligned}[t]
                                 & \exists W' \futurestr W \ldotp\exists \hs' \ldotp\\
                                 & \quad \heap' = \hs' \uplus \heap_f \land \\ 
                                 & \quad \heapSat[\hs']{n-i}{W'} \}
                               \end{aligned}
                             \end{aligned}
\end{align*}
